---
title: Homework- Week 5
author: Ayrianna Hedgecock
format: html
embed-resources: true
toc: true
---

## Problem 1


```{r}
#| label: Call libraries

library(flowCore)
library(flowWorkspace)
library(ggcyto)
library(ggplot2)
library(CytoML)
library(dplyr)
```


```{r}
#| label: File paths
#| error: true
#| eval: false
 
# DataFolder <- file.path("course" , "05_GatingSets" , "homeworks", "A-hedgecock-FlowCore", "data") # for actual code

DataFolder <- file.path("data") # Quarto Rendering
fcs_files <- list.files(DataFolder, pattern = ".fcs", full.names = TRUE)
FlowJoWspFiles <- list.files(DataFolder, pattern = ".wsp", full.names = TRUE)
OpenedWorkspace <- FlowJoWspFiles[stringr::str_detect(FlowJoWspFiles, "Opened")]
OpenedWorkspace
```

```{r}
#| eval: false
# hopefully skipping this block will allow rendering

ws <- open_flowjo_xml(file = OpenedWorkspace)
ws
```

```{r}
#| label: gating set

fj_gs <- flowjo_to_gatingset(
    ws = ws,
    name = 1,
    path = DataFolder,
    additional.keys = "GROUPNAME"
)
```


```{r}
#| label: gates

plot(fj_gs)
gateList <- gs_get_pop_paths(fj_gs, path = "auto")
counts <- gs_pop_get_count_fast(fj_gs)
```


```{r}
#| label: percentParent column

PercentParent <- mutate(
    counts,
    Percent_Parent = 
        round(Count / ParentCount, digits = 2)
)
```


```{r}
#| label: export stats practice
FinishedDataset <- paste0("PercentParentDataset", ".csv")
StorageLocation <- file.path(DataFolder, FinishedDataset)

write.csv(PercentParent, StorageLocation, row.names = FALSE)
```

## Problem 2

```{r}
#| label: Inital Gating Set- CytoML
#| error: true
GatingSetCytoML <- flowjo_to_gatingset(
    ws,
    name = 1,
    path = DataFolder
)
```

::: {.callout-warning title="Inital Error"} 

Error:
! Multiple FCS files match sample 00_Ctrl.fcs by filename, event count, and keywords.
Candidates are: 
/Users/ayrianna/Documents/Week5/A-hedgecock-FlowCore/data/2025_07_26_AB_02_INF052_00_Ctrl.fcs
/Users/ayrianna/Documents/Week5/A-hedgecock-FlowCore/data/2025_07_26_AB_02_INF100_00_Ctrl.fcs
/Users/ayrianna/Documents/Week5/A-hedgecock-FlowCore/data/2025_07_26_AB_02_INF179_00_Ctrl.fcs
Please move incorrect files out of this directory or its subdirectories.
:::


::: {.callout-warning title="New Error"}
Only these files in folder: 
    "A-hedgecock-FlowCore/data/2025_07_26_AB_02_INF179_00_Ctrl.fcs"
    "A-hedgecock-FlowCore/data/2025_07_26_AB_02_INF179_00_SEB.fcs"

Error:
! Duplicated GUIDs detected within group: 00_Ctrl.fcs_10000
 Consider adding additional keywords to the GUID with argument "additional.keys"or setting argument "additional.sampleID = TRUE" to disambiguate samples further.
:::

## Problem 3

Let's try FlowSet...

```{r}
fs <- read.flowSet(
    files = fcs_files,
    transformation = FALSE,
    truncate_max_range = FALSE
)
fs
```


```{r}
gs <- GatingSet(fs)

gs
```

This is just the fcs files so no gates should be present: 

```{r}
#| error: true
plot(gs)
```


```{r}
ggcyto(
    gs,
    subset = "root",
    aes(
        x = "TNFa",
        y = "IFNg"
    )
) + 
    geom_hex(bins = 128)
```

No data is showing?

Lets try just one file:


```{r}
ggcyto(
    gs[1],
    subset = "root",
    aes(
        x = "TNFa",
        y = "IFNg"
    )
) + 
    geom_hex(bins = 128)
```

Okay I see data here...

2 files?


```{r}
ggcyto(
    gs[c(1:2)],
    subset = "root",
    aes(
        x = "TNFa",
        y = "IFNg"
    )
) + 
    geom_hex(bins = 128)
```

Aha, actually data ia shown the whole time but very hard to see! The Scale is weird with the Ctrl vs SEB conditions...play around with the bins?


```{r}
ggcyto(
    gs,
    subset = "root",
    aes(
        x = "TNFa",
        y = "IFNg"
    )
) + 
    geom_hex(bins = 30)
```

Let's try on gated data:


```{r}
ggcyto(
    fj_gs,
    subset = "Tcells",
    aes(
        x = "TNFa",
        y = "IFNg"
    )
) + geom_hex(bins = 128)
```

Shows overlaid groups? how to sepearate by sample? I.e. sample = row, conditions = columns?

```{r}
ggcyto(
    fj_gs,
    subset = "Tcells",
    aes(
        x = "TNFa",
        y = "IFNg"
    )
) + geom_hex(bins = 128) +
    facet_grid(~name)
```


Not sure where to go from here... will prob learn more next week...